---
title: "Coding Assignment 2"
author: "Team 4"
date: "Due: 2022-10-23 23:59"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=TRUE}
#Put any packages you need here
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=TRUE}
# Put any packages you want here
library(gt)
library(tidyverse)
library(gtsummary)
library(plotly)
library(readxl)
library(plotly)
library(corrplot)
library(dplyr)
library(car)
library(jtools)
sessionInfo()
```

A Florida health insurance company wants to predict annual claims for individual clients. The company pulls a random sample of 50 customers. The owner wishes to charge an actuarially fair premium to ensure a normal rate of return. The owner collects all of their current customer’s health care expenses from the last year and compares them with what is known about each customer’s plan. 

The data on the 50 customers in the sample is as follows:

-	Charges: Total medical expenses for a particular insurance plan (in dollars)
-	Age: Age of the primary beneficiary
-	BMI: Primary beneficiary’s body mass index (kg/m2)
-	Female: Primary beneficiary’s birth sex (0 = Male, 1 = Female)
-	Children: Number of children covered by health insurance plan (includes other dependents as well)
-	Smoker: Indicator if primary beneficiary is a smoker (0 = non-smoker, 1 = smoker)
-	Cities: Dummy variables for each city with the default being Sanford

Answer the following questions using complete sentences and attach all output, plots, etc. within this report.


```{r dataset, include=FALSE}
insurance <- read_excel("../Data/insurance_0126_Group4.xlsx")
```



## Question 1

Generate this regression equation:

$$Charges = \beta_{0}+\beta_{1}*Age+\beta_{2}*BMI+\beta_{3}*Children + \beta_{4}*Female + \beta_{5}*Smoker + \beta_{6}*WinterSprings + \beta_{7}*WinterPark + \beta_{8}*Oviedo$$

```{r q1}
model_full <- lm(Charges ~., data = insurance)
summary(model_full)
```
$$Charges = -26631.15+235.89*Age+802.25*BMI+213.44*Children + 731.36 *Female + 20092.70*Smoker + 4109.84 *WinterSprings + 1962.97*WinterPark- 963.11*Oviedo$$
Which independent variables test significant in a two-tailed test? Use $\alpha = 0.2$
<br>
By analyzing the P values of each independent variable and considering that our alpha = 0.2, the independent variables that test significant are the following:<br>
WinterSprings with a P value of 0.110613 <br>
Smoker with a P value of 6.59e-12<br>
BMI with a P value of 2.96e-07 <br>
Age with a P value of 0.000746 <br>


## Question 2

Interpret the coefficients of all the independent variables that tested significant.
# Write relationship with the dependent variable and the significant independent variables

Age: For every year there is $236 added to the charges<br>
BMI: For every point increment in the BMI our charges go up in aprox. $802 <br>
Smoker: If the person is smoker it adds $20092.70 to the total charges <br>
WinterSprings: if the person lives in Winter Springs rather than Sanford the charges go up by $1962.97 <br>

## Question 3

The insurance agent returns with another client:

- Age: 35
- BMI: 26
- Children: 1
- Gender: Female
- Smoker: No
- City: Sanford

Using the regression equation above, predict the amount of medical expenses associated with this policy. (Provide a 95% confidence interval as well)

```{r q3}

newPrediction <- data.frame(Age=35, BMI=26, Children=1, Female=1, Smoker=0, WinterSprings=0, WinterPark=0, Oviedo=0)

predict(model_full, newdata= newPrediction, interval = "confidence")



```

Based on the prediction model the charges for an individual that is 35 years old, with a BMI of 26, non smoker, lives in Sanford, and has only one child will be aprox. $3,428, with the lowest amount being -1072.45 and the max. is $ 7928.80. Residual Standard Error is 5651 which indicates a poor model and allows there to be a negative number.


## Question 4

Compare the model from question 1, to a model that does not include the group of city dummies. Does including the cities reduce the SSE by a statistically significant amount? (check the partial F-test for significance of the dummies as a group). Use $\alpha = 0.05$

```{r q4}
nodummy <- insurance [,1:6, drop=FALSE]
nodummy <- lm (Charges~ ., data=nodummy)

anova(nodummy, model_full)

```

By comparing both models without the city variables showed no significant reduction SEE so there is no significant difference between one model that includes dummy variables and the second model that does not. With our P value being 0.2052 and our alpha = 0.05 we can make this conclusion.
